Discriminating 3D-objects using

### 2023-07-08 paper/paper.tex
merges

Agents interact with the physical world through their action and perception and with other agents through language.
Their sensors and actuators (if they are artificial agents but the same also holds for natural agents) allow them to sample the world and their their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
On the other hand language that is used to communicta ewith other agents is based on representations that are composed of a limited set of discrete symbols.
How can both domains and representations arising from these interaction be combined? How are ranges of measurements exprseed in a continuous domain mapped to discrete linguistic labels?
How is ambiguity and underpecification resolved?
Ho can agents achive this through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?

@@

% \section{Background and previous work}

Language games with a sender and a receiver \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa} offer an opportunity to examine how neural models condense seen information into a distinct representation of symbols through linguistic interaction.
A sender describes a perceptual situation that can also be seen by a receiver starting with a random label while
the receiver attempts to interpret the meaning of these symbols and combine it with other information.
It then sends feedback to the sender about its interpretation.
Interaction is optimised based on the communicative success of the sender and the receiver.
Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa,Kelleher:2020aa}.
Communication is successful and grounding converges because learning is constrained by joint attention and the agents are playing following the rule of the communicative games.

In this this paper we explore how agents based on artificial neural networks learn referential grounding of entitities in images of 3-dimensional scenes where one agent is discribing the entities and the other agent learns to interpret the reference of symbols in the scene either by identifying the object in the scene as a bounding box or location of the object based on object attributes such as shape, colour and size.
% studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object. Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that that can be referred to and discriminated based on the attributes such as shape, colour and size whereby the discrimination is based on different overalaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects aither as one of the potential bounding boxes or location of the target.

@@

\section{Our dataset: CLEVR-DaleTwo and DaleFive}

@@

The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{%The extended source code can be found on GitHub:
  \href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
This dataset includes 3D-generated images depicting scenes with different kinds of objects. Each of these objects has different combinations of attributes, such as \emph{shape}, \emph{color} and \emph{size}. The objects are randomly placed into the scene and assigned with random attributes.
Figure \ref{fig:clevr-extended_example} shows an example of a generated image of this CLEVR dataset. The dataset also contains ground truth information about each image, including attributes and locations for all objects.

\begin{figure}[h]
  \centering
  \frame{\includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}}
  \caption{Example of a generated image with the extended code} % SD 2023-07-08 15:00:20 +0200: What dataset: Dale-2?
  \label{fig:clevr-extended_example}
\end{figure}

@@

For this research, two datasets are created. The \emph{DaleTwo} dataset contains one target object and one distractor, while the \emph{DaleFive} dataset contains one target object and four distractors. Both datasets contain 10.000 images. For each image fixed size bounding boxes around each object are extracted. These bounding boxes are the input for the sender and receiver in the language game.

% SD 2023-07-08 15:23:38 +0200: Describe a bit more how DaleTwo and DaleFive differ, in particular how many attributes are shared and how between the objects in the DaleFive?

\subsection{Language game}
The language game was developed and run in the EGG framework \citep{Kharitonov2019}. The setup of the game is based on the game described in the paper by \citet{Lazaridou2016}. The same multiple images are passed to a sender and a receiver. The sender must communicate the target image to the receiver with a message, who needs to identify this target image. Instead of distinguishing two images from different concepts, in this experiment the agents need to distinguish two objects with shared attributes. Both sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} in \citet{Lazaridou2016}. One central difference is the production of the message. Their paper focuses on the classification of a concept for the input image and therefore produces only one-symbol messages that should correspond to these concepts. This research focuses on the identification of attributes for the objects and their combination, which is why our models produce a sequence of symbols for the message (which may correspond to the attributes). This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) (see Figure \ref*{todo}).

@@

The results will be evaluated using the accuracy if the receiver could identify the target object. A random guess corresponds to 50\% in the \emph{DaleTwo} dataset and 20\% in the \emph{DaleFive} dataset.

\section{Results}
TODO:
\begin{itemize}
  \item DaleTwo: \textbf{(done)}
        \begin{itemize}
          \item small hidden/embedding dims, small vocab -> high accuracy
          \item high hidden/embedding dims, small vocab -> low accuracy
        \end{itemize}
  \item DaleFive: \textbf{(done)}
        \begin{itemize}
          \item small hidden/embedding dims, small vocab -> low accuracy
          \item small hidden/embedding dims, bigger vocab -> higher accuracy
        \end{itemize}
  \item ... test different hidden/embedding dims
  \item ... test 3/4 objects
\end{itemize}

\begin{table}
  \centering
  \begin{tabular}{c|ccccc|c}
    \hline
    \textbf{Dataset} & $h_{s}$ & $e_{s}$ & $h_{r}$ & $e_{r}$ & $|V|$ & \textbf{Acc.} \\
    \hline
    DaleTwo          & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
    DaleTwo          & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
    DaleFive         & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
    DaleFive         & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
    DaleFive         & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
    \hline
  \end{tabular}
  \caption{Result of the experiments with different hidden sizes, embedding sizes and vocabulary sizes}
  \label{tab:results}
\end{table}

The results of the experiments are summarized in Table \ref{tab:results}. For the \emph{DaleTwo} dataset it can be clearly seen that very small embedding and hidden sizes are beneficial for identifying the correct object. The receiver identifies indeed almost every sample correctly with all sizes of 10. When the hidden and embedding sizes are increased, the guesses by the receiver are random. Interestingly, a vocab size of 10 is enough to communicate a meaningful message for the \emph{DaleTwo} dataset.

@@

The result change, when using the \emph{DateFive} dataset with four distractors. With low hidden, embedding and vocab size, the agents barely pass the random baseline with 23\%. Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\%. This accuracy is still far lower than the 95\% with the \emph{DaleTwo} dataset.

>>>>

Agents interact with the physical world through their action and perception and with other agents through language.
Their sensors and actuators (if they are artificial agents, but the same also holds for natural agents) allow them to sample the world and their own state using measures that are continuous in nature such as intensity of light, distance, angles, velocity and others which can be measured with a high degree of accuracy.
On the other hand language that is used to communicate with other agents is based on representations that are composed of a limited set of discrete symbols.
How can both domains and representations arising from these interactions be combined? How are ranges of measurements expressed in a continuous domain mapped to discrete linguistic labels?
How is ambiguity and underspecification resolved?
How can agents achieve this through interactive grounding \citep{Regier:1996,Roy:2005,Cooper:2023aa}?

@@

\section{Background and previous work}

Language games with a sender and a receiver \citep{Clark:1996aa,Bartlett:2005aa,Kirby:2008ab,SteelsLoetzsch:2009,Zaslavsky:2018aa} offer an opportunity to examine how neural models condense seen information into a distinct representation of symbols through linguistic interaction.
A sender describes a perceptual situation that can also be seen by a receiver starting with a random label while
the receiver attempts to interpret the meaning of these symbols and combine it with other information.
It then sends feedback to the sender about its interpretation.
Interaction is optimized based on the communicative success of the sender and the receiver.
Gradually, both agents converge on a set of symbols that they can use to refer to situations in jointly attended perceptual scenes \citep{Chai:2016aa, Kelleher2020}.
Communication is successful and grounding converges because learning is constrained by joint attention and the agents are playing following the rule of the communicative games.

In this paper we explore how agents based on artificial neural networks learn referential grounding of entities in images of 3-dimensional scenes where one agent is describing the entities and the other agent learns to interpret the reference of symbols in the scene either by identifying the object in the scene as a bounding box or location of the object based on object attributes such as shape, color and size.
% studies how the agents can encode real-looking objects into distinct symbols and decode this message to reidentify this object. Specifically, the study explores discrimination games using an extended version of the CLEVR dataset, focusing on the identification and differentiation of objects based on their attributes such as shape, color, and size.
The novelty of our work compared with the previous work with this setup \citep{Kharitonov2019} is that (i) we extend the popular CLEVR dataset \citep{Johnson2016} with new artificially generated 3-d scenes of objects that can be referred to and discriminated based on the attributes such as shape, color and size whereby the discrimination is based on different overlaps of these attributes between the target and the distractor objects; and (ii) we implement a focused referring to objects either as one of the potential bounding boxes or location of the target.

@@

\section{Our dataset: CLEVR-Dale-2 and Dale-5}

@@

The dataset used for these experiments is an extended version of the CLEVR dataset \citep{Johnson2016}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
This dataset includes 3D-generated images depicting scenes with different kinds of objects.
Each of these objects has different combinations of attributes, such as \emph{shape}, \emph{color} and \emph{size}.
The objects are randomly placed into the scene and assigned with random attributes.
Figure \ref{fig:clevr-extended_example} shows an example of a generated image of this CLEVR dataset.
The dataset also contains ground truth information about each image, including attributes and locations for all objects.

\begin{figure}[h]
  \centering
  \frame{\includegraphics[width=.8\linewidth]{figures/CLEVR_extended_example.png}}
  \caption{Example of a generated image in the Dale-5 dataset with the extended code}
  \label{fig:clevr-extended_example}
\end{figure}

@@

For this research, two datasets are created.
The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
Both datasets contain 10.000 images.
For each image fixed size bounding boxes around each object are extracted.
These bounding boxes are the input for the sender and receiver in the language game.

% SD 2023-07-08 15:23:38 +0200: Describe a bit more how Dale-2 and Dale-5 differ, in particular how many attributes are shared and how between the objects in the Dale-5?

\section{Language game}
The language game was developed and run in the EGG framework \citep{Kharitonov2019}.
The setup of the game is based on the game described in the paper by \citet{Lazaridou2016}.
The same multiple images are passed to a sender and a receiver.
The sender must communicate the target image to the receiver with a message, who needs to identify this target image.
Instead of distinguishing two images from different concepts, in this experiment the agents need to distinguish two objects with shared attributes.
Both sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} in \citet{Lazaridou2016}.
One central difference is the production of the message.
Their paper focuses on the classification of a concept for the input image and therefore produces only one-symbol messages that should correspond to these concepts.
This research focuses on the identification of attributes for the objects and their combination, which is why our models produce a sequence of symbols for the message (which may correspond to the attributes).
This is done using an encoder LSTM (sender) and a decoder LSTM (receiver) (see Figure \ref*{todo}).

@@

The results will be evaluated using the accuracy if the receiver could identify the target object.
A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.

\section{Results}
TODO:
\begin{itemize}
  \item Dale-2: \textbf{(done)}
        \begin{itemize}
          \item small hidden/embedding dims, small vocab -> high accuracy
          \item high hidden/embedding dims, small vocab -> low accuracy
        \end{itemize}
  \item Dale-5: \textbf{(done)}
        \begin{itemize}
          \item small hidden/embedding dims, small vocab -> low accuracy
          \item small hidden/embedding dims, bigger vocab -> higher accuracy
        \end{itemize}
  \item ... test different hidden/embedding dims
  \item ... test 3/4 objects
\end{itemize}

\begin{table}
  \centering
  \begin{tabular}{c|ccccc|c}
    \hline
    \textbf{Dataset} & $h_{s}$ & $e_{s}$ & $h_{r}$ & $e_{r}$ & $|V|$ & \textbf{Acc.} \\
    \hline
    Dale-2           & {10}    & {10}    & {10}    & {10}    & {10}  & {95\%}        \\
    Dale-2           & {50}    & {50}    & {128}   & {128}   & {10}  & {50\%}        \\
    Dale-5           & {10}    & {10}    & {10}    & {10}    & {10}  & {23\%}        \\
    Dale-5           & {10}    & {10}    & {10}    & {10}    & {20}  & {23\%}        \\
    Dale-5           & {10}    & {10}    & {10}    & {10}    & {100} & {41\%}        \\
    \hline
  \end{tabular}
  \caption{Result of the experiments with different hidden sizes, embedding sizes and vocabulary sizes}
  \label{tab:results}
\end{table}

The results of the experiments are summarized in Table \ref{tab:results}.
For the \emph{Dale-2} dataset it can be clearly seen that very small embedding and hidden sizes are beneficial for identifying the correct object.
The receiver identifies indeed almost every sample correctly with all sizes of 10.
When the hidden and embedding sizes are increased, the guesses by the receiver are random.
Interestingly, a vocab size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.

@@

The result change, when using the \emph{Date-5} dataset with four distractors.
With low hidden, embedding and vocab size, the agents barely pass the random baseline with 23\%.
Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\%.
This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.
